Coherent IR-Hyper-Raman Four Wave Mixing Vibrational Spectroscopy

### 2024-07-29 paper.tex
Added F factor reference and definition in paper

It is useful to expose relationships between nonlinear output, transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit. \cite{Simpson2004}
Under the electric dipole approximation, the I$^{th}$ component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation) \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(\omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the IJKL element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}
	\chi^{(3)}_{IJKL} = NF(\omega_4) \langle \gamma_{ijkl} \rangle
\end{equation}
where $N$ is a number density, $F$ is the Lorentz local field factor, and $\gamma_{ijkl}$ is the third order polarizability (i.e., second hyperpolarizability). 
The brackets indicate an orientational (or rotational) average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.

@@

\subsection{Time Ordering}
It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.
While we have only isolated one specific HDFG process (\autoref{fig:comparisonwmel}), there are at least fourteen different WMELs which can provide HDFG output in the driven limit. \cite{RN352}
If $\omega_2$ ($\omega_1$) becomes resonant (non-resonant), where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the WMEL diagrams presented in \autoref{fig:sivewmel2} will interfere, as they share the same phasematching condition but have oppositely signed amplitudes.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66 in]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$.
		In all cases, $\omega_2 > \omega_1$.
	}
	\label{fig:sivewmel2}
\end{figure*}

@@

Despite their increasing use for dissecting intra and intermolecular interactions, four wave mixing techniques are usually not feasible for most laboratories, as they usually demand the use of multiple optical parametric amplifiers and/or complex acousto-optic modulators. \cite{Chen2016}
Unlike most four wave mixing techniques, HDFG can be performed using only two input beams (i.e., two color HDFG).
In the case of two color HDFG, only one tunable infrared optical parametric amplifier is needed, as the oscillator which pumps the infrared OPA can be used to perform the two-photon upconversion needed to generate HDFG output, i.e., $-\vec{k}_1 + 2\vec{k}_2$.
Vibrational sum frequency generation (vSFG) is a $\chi^{(2)}$ technique whose output, when only vibrationally resonant, scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor, whose phasematching constraint is  $\vec{k}_1 + \vec{k}_2$.
Laboratories which use vSFG to investigate interfacial species at buried interfaces thus have the ingredients necessary to perform a HDFG experiment in a transmission geometry. \cite{Piontek2023_1}
This would allow vSFG practitioners to perform measurements in the bulk, such as measuring free induction decay, vibrational population lifetimes, and steady-state vibrational spectra through HDFG.

@@

In HDFG, 
\begin{equation}\label{chi3}
\begin{split}
		\chi^{(3)}_{IJKL} &= NF(\omega_4) \langle \gamma_{ijkl} \rangle = -\frac{NF}{4D \hbar \varepsilon_0 \Delta_{gv}} \langle \beta_{ijk} \mu_l \rangle \rho_{gg}\\
\end{split}
\end{equation}
For simplicity, we take $\rho_{gg} = 1$.
The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Briefly, a tensor in the molecular frame, A$_{ijkl}$, is transformed into an element of the same tensor in the laboratory frame, A$_{IJKL}$, through A$_{IJKL}$ = $\theta^{ijkl}_{IJKL} A_{ijkl} = \langle A_{ijkl} \rangle$, where summation over repeated indices is implied and $\theta$ is the transformation operator. \cite{McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. \cite{Bersohn1966, Willetts1992, Kauranen1996}
By using the expansions of $\beta_{ijk}$ and $\mu_{l}$ to $\order{Q_v}$ found earlier, and knowing that $\Re(\chi^{(3)})$ vanishes when resonant, we see
\begin{equation}\label{betasive}
	\left\langle \frac{\partial \beta_{ijk}}{\partial Q_v} {\frac{\partial \mu_l}{\partial Q_v}} \right\rangle = -\frac{8D \varepsilon_0}{NF}  {\Gamma_{gv} \omega_{vg}} {\Im(\chi^{(3)}_{IJKL})}
\end{equation}
Since $\abs{\partial \mu / \partial Q}$ values can be extracted from FT-IR spectra,\cite{RN459} HDFG can give quantitative information on the magnitude of $\beta_{ijk}$ for infrared active vibrations.
Additionally, by using non-degenerate input frequencies to stimulate the hyper-Raman transition, the asymmetric properties of $\beta_{ijk}$ can be examined and quantified. \cite{Christie1971, Denisov1986, Kozich2007}
These quantitative aspects of HDFG could prove useful for examining how well computational methods calculate $\beta_{ijk}$ values and the strength of spontaneous hyper-Raman scattering with non-degenerate pulses.

@@

\section{Appendix: Comparison of vSFG and two-beam HDFG output}\label{appendixA}
 To motivate the application of HDFG spectroscopy, we perform a calculation to compare two-color HDFG and vSFG output, where absorption effects are neglected for simplicity.
 The output polarizations under the electric dipole approximation are written as

	 	\begin{subequations}
		 	\begin{equation}\label{PSFG}
		 		\begin{split}
		 		\abs{P^{(2)}_{\text{vSFG}}} &= \frac{N_\text{surf}}{\ell} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E(\omega_2)E(\omega_1)} 
		 		\end{split}
			 \end{equation}
		 	\begin{equation}
		 		\begin{split}
			 		\abs{P^{(3)}_{\text{HDFG}}} &= N_\text{bulk}  F(2\omega_2-\omega_1)  \abs{\langle \beta_{ijk} \mu_{l} \rangle E{(\omega_2)}E(\omega_2)E(-\omega_1)}
		 		\end{split}
			 \end{equation}
		 \end{subequations}

 where N$_\text{bulk}$ is the bulk number density ($\sim$ 10$^{28}$ m$^{-3}$ for H$_2$O$_{(l)}$ at room temperature), and N$_\text{surf}$ is the surface number density ($\sim$ 10$^{16}$ m$^{-2}$ for H$_2$O adsorbed at quartz).\cite{Du1994}	
 For simplicity, we assume that $\chi^{(2)}$ is a spatial constant so that $\int_{0^-}^\ell \mathrm{d}z \langle \alpha_{ij}\mu_{k} \rangle \approx \ell \langle \alpha_{ij}\mu_{k} \rangle$, where $z$ is oriented along the surface normal and $0^{-}$ is the edge of the interface where substrate begins. 
 We take the film thickness ($\ell$) probed by vSFG to be 1 nm, much less than the vSFG coherence length.\cite{RN133}
 A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(2\omega_2-\omega_1)$.
 To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
 \begin{equation}
	 		P_\text{ratio} \equiv \frac{\abs{P^{(3)}_{\text{HDFG}}}}{\abs{P^{(2)}_{\text{vSFG}}}} \approx \frac{N_\text{bulk}}{N_\text{surf}} \frac{\beta}{\alpha} E(\omega_2) \ell \sim 10^3 \frac{\beta}{\alpha} E(\omega_2)\\
	 \end{equation}
 Ziegler has noted that for a field with intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_2)$ is largely detuned from electronic resonances. \cite{RN515}
 Such an intensity is easily obtained using modern ultrafast sources.
 In this limit, $P_\text{ratio} \sim 1$.
 Since the intensity ratio scales as $\abs{P_\text{ratio}}^2$, we see that the HDFG output is roughly as strong as vSFG, assuming only interfacial contributions in vSFG.
 It is important to note that this calculation assumed negligible dipole and Raman polarizability dependence upon the distance from surface normal, ignored orientational averaging effects, and presumed equivalence of local field factors, all of which can significantly reduce output from either process. 
 Nevertheless, since HDFG produces a number of photons similar to that of vSFG, HDFG is a viable technique for investigating bulk systems.
 This method provides practitioners of sum frequency generation the ability to probe isotropic dephasing dynamics without many meaningful changes in their optical setup, and thus compare  dephasing in the bulk and at the interface.\cite{RN224}

>>>>

It is useful to expose relationships between nonlinear output, transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit. \cite{Simpson2004}
Under the electric dipole approximation, the I$^{th}$ component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation) \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(\omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the IJKL element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}
	\chi^{(3)}_{IJKL} = NF \langle \gamma_{ijkl} \rangle
\end{equation}
where $N$ is a number density, $F$ is the Lorentz local field factor, and $\gamma_{ijkl}$ is the third order polarizability (i.e., second hyperpolarizability). 
For $j$ input and output fields with frequency $\omega_j$, 
\begin{equation}
	F = \prod_j \frac{\left(n(\omega_j)\right)^2 + 1}{3}
\end{equation}
where $n(\omega_j)$ is the refractive index at frequency $\omega_j$.\cite{Bedeaux1973}
The brackets indicate an orientational (or rotational) average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.

@@

\subsection{Time Ordering}
It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.
While we have only isolated one specific HDFG process (\autoref{fig:comparisonwmel}), there are at least fourteen different WMELs which can provide HDFG output in the driven limit. \cite{RN352}
If $\omega_2$ ($\omega_1$) becomes resonant (non-resonant), where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the WMEL diagrams presented in \autoref{fig:sivewmel2} will interfere, as they share the same phase-matching condition but have oppositely signed amplitudes.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66 in]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$.
		In all cases, $\omega_2 > \omega_1$.
	}
	\label{fig:sivewmel2}
\end{figure*}

@@

Despite their increasing use for dissecting intra and intermolecular interactions, four wave mixing techniques are usually not feasible for most laboratories, as they usually demand the use of multiple optical parametric amplifiers and/or complex acousto-optic modulators. \cite{Chen2016}
Unlike most four wave mixing techniques, HDFG can be performed using only two input beams (i.e., two color HDFG).
In the case of two color HDFG, only one tunable infrared optical parametric amplifier is needed, as the oscillator which pumps the infrared OPA can be used to perform the two-photon upconversion needed to generate HDFG output.
Vibrational sum frequency generation (vSFG) is a $\chi^{(2)}$ technique whose output, when only vibrationally resonant, scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor, whose phase-matching constraint is  $\vec{k}_3 = \vec{k}_1 + \vec{k}_2$.
HDFG is phase-matchable under $\vec{k}_4 = -\vec{k}_1 + 2\vec{k}_2$ conditions for most, if not all, vibrational resonances.
Laboratories which use vSFG to investigate interfacial species at buried interfaces thus have the ingredients necessary to perform a HDFG experiment in a transmission geometry. \cite{Piontek2023_1}
This would allow vSFG practitioners to perform measurements in the bulk, such as measuring steady-state vibrational spectra, free induction decay and vibrational population lifetimes through HDFG or pump-HDFG-probe.

@@

In HDFG, 
\begin{equation}\label{chi3}
\begin{split}
		\chi^{(3)}_{IJKL} &= NF \langle \gamma_{ijkl} \rangle = -\frac{NF}{4D \hbar \varepsilon_0 \Delta_{gv}} \langle \beta_{ijk} \mu_l \rangle \rho_{gg}\\
\end{split}
\end{equation}
For simplicity, we take $\rho_{gg} = 1$.
The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Briefly, a tensor in the molecular frame, A$_{ijkl}$, is transformed into an element of the same tensor in the laboratory frame, A$_{IJKL}$, through A$_{IJKL}$ = $\theta^{ijkl}_{IJKL} A_{ijkl} = \langle A_{ijkl} \rangle$, where summation over repeated indices is implied and $\theta$ is the transformation operator. \cite{McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. \cite{Bersohn1966, Willetts1992, Kauranen1996}
By using the expansions of $\beta_{ijk}$ and $\mu_{l}$ to $\order{Q_v}$ found earlier, and knowing that $\Re(\chi^{(3)})$ vanishes when resonant, we see
\begin{equation}\label{betasive}
	\left\langle \frac{\partial \beta_{ijk}}{\partial Q_v} {\frac{\partial \mu_l}{\partial Q_v}} \right\rangle = -\frac{8D \varepsilon_0}{NF}  {\Gamma_{gv} \omega_{vg}} {\Im(\chi^{(3)}_{IJKL})}
\end{equation}
Since $\abs{\partial \mu / \partial Q}$ values can be extracted from FT-IR spectra,\cite{RN459} HDFG can give quantitative information on the magnitude of $\beta_{ijk}$ for infrared active vibrations.
Additionally, by using non-degenerate input frequencies to stimulate the hyper-Raman transition, the asymmetric properties of $\beta_{ijk}$ can be examined and quantified. \cite{Christie1971, Denisov1986, Kozich2007}
These quantitative aspects of HDFG could prove useful for examining how well computational methods calculate $\beta_{ijk}$ values and the strength of spontaneous hyper-Raman scattering with non-degenerate pulses.

@@

\section{Appendix: Comparison of vSFG and two-beam HDFG output}\label{appendixA}
 To motivate the application of HDFG spectroscopy, we perform a calculation to compare two-color HDFG and vSFG output, where absorption effects are neglected for simplicity. \cite{Carlson1989}
 The output polarizations under the electric dipole approximation are written as
	 	\begin{subequations}
		 	\begin{equation}\label{PSFG}
		 		\begin{split}
		 		\abs{P^{(2)}_{\text{vSFG}}} &= \frac{N_\text{surf}}{\ell} \abs{F_{\text{SFG}} \langle \alpha_{ij}\mu_{k} \rangle E(\omega_2)E(\omega_1)} 
		 		\end{split}
			 \end{equation}
		 	\begin{equation}
		 		\begin{split}
			 		\abs{P^{(3)}_{\text{HDFG}}} &= N_\text{bulk} \abs{F_{\text{HDFG}} \langle \beta_{ijk} \mu_{l} \rangle E{(\omega_2)}E(\omega_2)E(-\omega_1)}
		 		\end{split}
			 \end{equation}
		 \end{subequations}
 where N$_\text{bulk}$ is the bulk number density ($\sim$ 10$^{28}$ m$^{-3}$ for H$_2$O$_{(l)}$ at room temperature), and N$_\text{surf}$ is the surface number density ($\sim$ 10$^{16}$ m$^{-2}$ for H$_2$O adsorbed at quartz).\cite{Du1994}	
 For simplicity, we assume that $\chi^{(2)}$ is a spatial constant so that $\int_{0^-}^\ell \mathrm{d}z \langle \alpha_{ij}\mu_{k} \rangle \approx \ell \langle \alpha_{ij}\mu_{k} \rangle$, where $z$ is oriented along the surface normal and $0^{-}$ is the edge of the interface where substrate begins. 
 We take the film thickness ($\ell$) probed by vSFG to be 1 nm, much less than the vSFG coherence length.\cite{RN133}
 A normal dispersion curve is assumed so that $n(\omega_1+\omega_2) \approx n(2\omega_2-\omega_1)$.
 To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
 \begin{equation}
	 		P_\text{ratio} \equiv \frac{\abs{P^{(3)}_{\text{HDFG}}}}{\abs{P^{(2)}_{\text{vSFG}}}} \approx \frac{N_\text{bulk}}{N_\text{surf}} \abs{\frac{\left(n(\omega_2)\right)^2 + 1}{3}} \frac{\beta}{\alpha} E(\omega_2) \ell \sim 10^3 \frac{\beta}{\alpha} E(\omega_2)\\
	 \end{equation}
 Ziegler has noted that for a field with intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_2)$ is largely detuned from electronic resonances. \cite{RN515}
 Such an intensity is easily obtained using modern ultrafast sources.
 In this limit, $P_\text{ratio} \sim 1$.
 Since the intensity ratio scales as $\abs{P_\text{ratio}}^2$, we see that the HDFG output is roughly as strong as vSFG, assuming only interfacial contributions in vSFG.
 It is important to note that this calculation assumed negligible dipole and Raman polarizability dependence upon the distance from surface normal, ignored orientational averaging effects, and presumed equivalence of local field factors, all of which can significantly reduce output from either process. 
 Nevertheless, since HDFG produces a number of photons similar to that of vSFG, HDFG is a viable technique for investigating bulk systems.
 This method provides practitioners of sum frequency generation the ability to probe isotropic dephasing dynamics without many meaningful changes in their optical setup, and thus compare  dephasing in the bulk and at the interface.\cite{RN224}
